### 2018-09-17 algorithm.tex
fix #14: calibrate US wages from trend deviations from value added share

Because in expectation the last term on the right-hand-side of this equation is zero, the $L^*$ has to be such that
\[
E_{A}\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
= \lambda L_{nt}
\]
constant. There can be no expected deviations in the (ratio) relative wage across sectors. 
\begin{equation}\label{eq:varepsilon}
\frac {\varepsilon_{njt}} {L_{nt}}
=
\frac 1\varrho
\left[
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	-\lambda L_{nt}
\right],
\end{equation}
and $\lambda$ is such that $\sum_j\varepsilon_{njt}=0$ because of the resource constraint.

@@

Assume that the US is in a flexible-labor equilibrium in every period so that 
\[
w_{njt} = w_{nt},
\]
nominal GDP. 
\end{document}

>>>>

Because in expectation the last term on the right-hand-side of this equation is zero, the $L^*$ has to be such that
\[
E_{A}\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
= \lambda L_{nt}
\]
constant. There can be no expected deviations in the (ratio) relative wage across sectors. 
\begin{equation}\label{eq:varepsilon}
\frac {\varepsilon_{njt}} {L_{nt}}
=
\frac 1\varrho
\left[
\frac 
	{w_{njt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	{w_{nt}(\mathbf A_{t}, \mathbf L_{t}^*+\varepsilon_t)}
	-\lambda L_{nt}
\right],
\end{equation}
and $\lambda$ is such that $\sum_j\varepsilon_{njt}=0$ because of the resource constraint. That is, 
\[
\lambda L_{nt} = \frac1J \sum_{j=1}^{J} \frac{w_{njt}} {w_n},
\]
the average wage ratio across sectors (which may be different from one due to Jensen inequality terms).

@@

Using \eqref{eq:varepsilon} to express sectoral wages in the US,
\[
w_{njt} = w_{nt}
\left[
\frac {\varrho(L_{njt}-L_{njt}^*)}
	{L_{nt}}
+\lambda L_{nt}
\right].
\]
How do we calibrate this without observing $L_{njt}$? Multiply by $w_{njt}/w_{nt}$,
\[
\left(\frac {w_{njt}} {w_{nt}}\right)^2
= 
\varrho\frac {w_{njt}L_{njt}-w_{njt}L_{njt}^*}
	{w_{nt}L_{nt}}
+\lambda L_{nt}\frac {w_{njt}} {w_{nt}}
=\varrho (V_{njt} - V_{njt}^*)
+\lambda L_{nt}\frac {w_{njt}} {w_{nt}}.
\]
The first term is value added share at current and at expected allocations, denoted $V_{njt}$ and $V_{njt}^*$, respectively. Solving this quadratic equation,
\[
\frac {w_{njt}} {w_{nt}} = \frac12 \lambda L_{nt}
+\frac12 \sqrt{\lambda^2 L_{nt}^2 + 4\varrho (V_{njt}-V_{njt}^*)}.
\]
Recall that $\lambda L_{nt}$ equals the unweighted average wage ratio across sectors, and is approximately one, ignoring Jensen inequality terms. This helps us calibrate the sectoral wage rate as
\[
w_{njt} \approx w_{nt} \left[
\frac12
+\frac12 \sqrt{1 + 4\varrho (V_{njt}-V_{njt}^*)}
\right].
\]
We measure $V_{njt}$ as value added share in the data, and approximate $V_{njt}^*$ as the trend of value added shares.

\end{document}
